---
title: "5205 HW3 Q5 clean-up"
author: "Peiru Wu"
date: "October 5, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


my solution is as following:
(a)
```{r}
CPI <- read.csv("/Users/peiruwu/Downloads/cpiu-long.csv",header = T)
class(CPI)
summary(CPI)
```
(b)
```{r}
cpi.vec.ym <- CPI$X.2[!is.na(CPI$X.2)]
length(cpi.vec.ym)
year.month.vec <- CPI$X[!is.na(CPI$X.2)]
length(year.month.vec)
```
(c)
```{r}

plot(year.month.vec,cpi.vec.ym)
m<-length(cpi.vec.ym)
year.month.vec

cpi.vec<-c()
year.vec<-c(1913:2016)

for (i in 1913:2016){
  cpi.temp<-c()
  for(j in 1:m){
    if (year.month.vec[j]==i){
    cpi.temp<-append(cpi.temp,cpi.vec.ym[j])
  }}
    cpi.vec<-append(cpi.vec,mean(cpi.temp))
}
plot(year.vec,cpi.vec)
```
(d)
```{r}
cpi.vec.log<-log(cpi.vec)
plot(year.vec,cpi.vec.log)
##(we do not want inflation to raise because as inflation raises, the same amount of salary will value less.)
summary(lm(cpi.vec.log~year.vec))
## I suggest linear model.my beta0 is -6.123e+01 and my beta1 is 3.311e-02.
abline(lm(cpi.vec.log~year.vec),year.vec)
##I think the fit is good.year about 1975,2009 deviate a little bit more.
```
(e)
```{r}
predict(lm(cpi.vec.log~year.vec),data.frame(year.vec=c(2017,2026)))
min.raise.2017<-100*(exp(5.545135)-cpi.vec[104])/cpi.vec[104]
min.raise.2017    
min.raise.2026<-100*(exp(5.843086)-cpi.vec[104])/cpi.vec[104]
min.raise.2026   
```
(f)
in 1929, the cpi is higher than we thought because of inflation bubble.the same as 2009. 
(g)
```{r}
cpi.vec.log2<-cpi.vec.log[-c(1:28)]  # exclude the data poins of the year 1929 to 1940
year.vec2<-year.vec[-c(1:28)]
plot(year.vec2,cpi.vec.log2)
summary(lm(cpi.vec.log2~year.vec2))
abline(lm(cpi.vec.log2~year.vec2),year.vec2)

predict(lm(cpi.vec.log2~year.vec2),data.frame(year.vec2=c(2017,2026)))
min.raise2.2017<-100*(exp(5.75868)-cpi.vec[104])/cpi.vec[104]
min.raise2.2017  
min.raise2.2026<-100*(exp(6.11831)-cpi.vec[104])/cpi.vec[104]
min.raise2.2026   



